---
title: "Lot.size_ALPS"
author: "ALPS"
date: "27-03-2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Starting: loading packages and data
```{r message=FALSE, warning=FALSE, paged.print=TRUE}
rm(list=ls())
set.seed(13)

# # # # # # # # # # # # # # # # # # # # # # # # 
# Models testing # LOTTIA SIZE
# # # # # # # # # # # # # # # # # # # # # # # #

# load packages
library(lme4)
library(bbmle)
library(glmmTMB)
library(tidyverse)
library(nlme)
library(DHARMa)
library(lmerTest)
library(nlme)
library(lattice)
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)

source("functions/HighstatLibV6.r")

# # # LOAD DATA
lapas_sz1 <- read.csv("data/lottia_sz2020.csv", header=T)
nodi_sz <- read.csv("data/nodi_sz2020.csv", header=T)

ordem <- unique(nodi_sz[order(nodi_sz$distance_S),]['site'])
lapas_sz1$site <- factor(lapas_sz1$site, levels = as.character(ordem$site))
lapas_sz1$subregion <- factor(lapas_sz1$subregion, levels = c("MRBS","SSCh","Ubatuba","SCRJ","MRRJ","LRRJ"))

lapas_sz <- lapas_sz1 %>% 
  filter(estagio_recruta_adulto == 'A') %>% 
  group_by(site, quadrado) %>% 
  summarise(mean_size = mean(tamanho_mm)) %>% 
  dplyr::select(site, mean_size) %>% 
  inner_join(., unique(lapas_sz1[,c(1,6:10,12:18)]), by='site') %>% 
  data.frame()

rm(lapas_sz1, nodi_sz)

hist(lapas_sz$mean_size)
shapiro.test(lapas_sz$mean_size)
hist(log(lapas_sz$mean_size))
shapiro.test(log(lapas_sz$mean_size))

hist(log(lapas_sz$mean_size))## log-transformation (log (x)) improves normal distribution (USE it) ##
```


#### Plotting for exploration
```{r data exploration}
ggplot(lapas_sz, aes(mean_size)) +
  geom_density(aes(fill=site), alpha=0.5) +
  #geom_histogram(aes(fill=site)) +
  facet_grid(subregion~., drop = TRUE) +
  theme_classic() +
  theme(legend.position = "")


ggplot(lapas_sz, aes(x = site, y = mean_size, color = subregion)) + 
  geom_boxplot(notch=F) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5))
```

#### VIF and correlations
```{r VIF}
lapas_sz %>% 
  na.omit() %>% 
  dplyr::select(-region,-subregion,-site,-easting,-northing) %>% 
  corvif()

lapas_sz %>% 
  dplyr::select(-region,-subregion,-site) %>% 
  PerformanceAnalytics::chart.Correlation(histogram=TRUE, pch=19)

# visual inspection
lapas_sz %>% 
  dplyr::select(-region,-subregion, -ext_m_tetra) %>% 
  group_by(site) %>% 
  summarise_if(is.numeric, mean, na.rm = TRUE) %>% 
  gather(key='fatores', value='valores', -site, -mean_size) %>% 
  ggplot() + 
  geom_point(aes(x= valores, y=mean_size)) + 
  geom_smooth(aes(x= valores, y=mean_size)) + 
  facet_wrap(~ fatores, ncol=3, scales = 'free') +
  theme_classic()
```


#### Selecting random structure of predictive model
```{r dandom, Distribution, message=FALSE, warning=FALSE}
size.env.full = glmmTMB(log(mean_size) ~ ndvi + sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + (1|region/subregion/site), data=lapas_sz, REML=T)

size.env.no_reg = glmmTMB(log(mean_size) ~ ndvi +  sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + (1|subregion/site), data=lapas_sz, REML=T)

size.env.no_sub = glmmTMB(log(mean_size) ~ ndvi +  sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + (1|region/site), data=lapas_sz, REML=T)

size.env.no_loc = glmmTMB(log(mean_size) ~ ndvi +  sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + (1|region/subregion), data=lapas_sz, REML=T)

size.env.reg = glmmTMB(log(mean_size) ~ ndvi + sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + (1|region), data=lapas_sz, REML=T)

size.env.sub = glmmTMB(log(mean_size) ~ ndvi + sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + (1|subregion), data=lapas_sz, REML=T)

size.env.loc = glmmTMB(log(mean_size) ~ ndvi + sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + (1|site), data=lapas_sz, REML=T)

cand.set.env <- list(size.env.full, size.env.no_reg, size.env.no_sub, size.env.no_loc, size.env.reg, size.env.sub, size.env.loc)

modnames.env<-c("size.env.full","size.env.no_reg","size.env.no_sub","size.env.no_loc", "size.env.reg", "size.env.sub", "size.env.loc")

AICctab(cand.set.env, mnames=modnames.env, base=TRUE, weights=TRUE, logLik=TRUE)

anova(size.env.no_reg, size.env.full)

## no significant difference, model dens.g1.no_reg is simpler ##
## CONCLUSION: best random structure includes (1|subregion/site) ##
```

#### Selecting fixed structure of predictive model (method = ML)  
```{r fixed}
dens.g1.ML = glmmTMB(log(mean_size) ~ ndvi + chl_mean + sst_mean + fwd_mean + wf_log + rugosity +                  incl_graus_tetra +(1|subregion/site), data=lapas_sz, REML=FALSE)
summary(dens.g1.ML)

dens.g2.ML <- update(dens.g1.ML, .~. -ndvi)
summary(dens.g2.ML)

dens.g3.ML <- update(dens.g2.ML, .~. -chl_mean)
summary(dens.g3.ML) 

dens.g4.ML <- update(dens.g3.ML, .~. -incl_graus_tetra)
summary(dens.g4.ML) 

dens.g5.ML <- update(dens.g4.ML, .~. -wf_log)
summary(dens.g5.ML) 

anova(dens.g4.ML, dens.g5.ML) 
## there is no difference between model with or without wve fetch; select simpler one ##
## CONCLUSION: selected model includes: Chla, SST, fwd and rugosity ##
```

#### Refit with REML and summary 
```{r REML}
dens.g.REML <- glmmTMB(log(mean_size) ~ sst_mean + fwd_mean + rugosity + (1|subregion/site), data=lapas_sz, REML=T) 
summary(dens.g.REML)
```

#### Checking model assumptions 
```{r residuals}
res.lot1 = xyplot(resid(dens.g.REML, type="pearson") ~ fitted(dens.g.REML), 
       ylab="Pearson residuals", xlab="Fitted values", panel = function(x,y){
         panel.grid(h = -1, v = 2)
         panel.points(x, y, col = 1)
         panel.loess(x, y, span = 0.5, col = 1,lwd=2)})

res.lot2 = xyplot(resid(dens.g.REML, type="pearson") ~ sst_mean, data=lapas_sz, 
       ylab="Pearson residuals", xlab="SST mean", panel = function(x,y){
         panel.grid(h = -1, v = 2)
         panel.points(x, y, col = 1)
         panel.loess(x, y, span = 0.5, col = 1,lwd=2)})

res.lot3 = xyplot(resid(dens.g.REML, type="pearson") ~ fwd_mean, data=lapas_sz, 
       ylab="Pearson residuals", xlab="fwd mean", panel = function(x,y){
         panel.grid(h = -1, v = 2)
         panel.points(x, y, col = 1)
         panel.loess(x, y, span = 0.5, col = 1,lwd=2)})

res.lot4 = xyplot(resid(dens.g.REML, type="pearson") ~ rugosity, data=lapas_sz, 
       ylab="Pearson residuals", xlab="rugosity index", panel = function(x,y){
         panel.grid(h = -1, v = 2)
         panel.points(x, y, col = 1)
         panel.loess(x, y, span = 0.5, col = 1,lwd=2)})

res.lot5 = histogram(resid(dens.g.REML, type="pearson"), col="white", xlab=list(label="Pearson residuals"))

lottia.residuals = grid.arrange(res.lot1, res.lot2, res.lot3, res.lot4, res.lot5, ncol=2) ### plotting together ###
plot(lottia.residuals)

## CONCLUSION: there are problems on residuals (apparently) ##

####
#lapas_sz %>% 
  select(ndvi, chl_mean, sst_mean, fwd_mean, wf_log, rugosity, incl_graus_tetra, mean_size) %>% 
  gather(key='fatores', value='valores', -mean_size) %>% 
  bind_cols(resid=rep(resid(dens.g.REML,  type="pearson"), 7)) %>% 
  ggplot() + 
  geom_point(aes(x= valores, y=resid)) + 
  geom_smooth(aes(x= valores, y=resid)) +
  facet_wrap(~ fatores, ncol=3, scales = 'free') +
  theme_classic()
```

#### Plotting to check spatial autocorrelation
```{r check spatial}
lapas_sz2= as.data.frame(lapas_sz)
##create jitter (values a little different) because correlation term doesn't work when distance is 0 between replicates###
lapas_sz2$easting2 =  jitter(lapas_sz2$easting)
lapas_sz2$northing2 = jitter(lapas_sz2$northing)

library(gstat)
library(sp)
library(ncf)

## getting the residuals ##
E <- residuals(dens.g.REML, type="pearson") 
## creating dataframe with residuals and coordinates ##
mydata <- data.frame(E, lapas_sz$easting, lapas_sz$northing) 
## creating spatial object ##
coordinates(mydata) <- c("lapas_sz.easting","lapas_sz.northing") 
## plotting bubblemap with residuals ##
bubble(mydata, "E", col = c("black","red"), alpha=0.1,
       main = "Pearson Residuals", xlab = "Northing",
       ylab = "Easting", scales = list(draw = T)) 

vario1 <- variogram(E ~ 1, mydata)
plot(vario1)## there is apparently correlation in the residuals at distance 10000 ##

## CONCLUSION: apparently there are not spatial patterns on residuals ##
```

#### Testing for spatial autocorrelation 
```{r spatial model}
# dens.g.REML <- glmmTMB(log(mean_size) ~ sst_mean + fwd_mean + rugosity + (1|subregion/site),                         data=lapas_sz, REML=T) 
# 
# mod.gau = lme(log(mean_size) ~ sst_mean + fwd_mean + rugosity, random=~ 1|subregion/site,         data=lapas_sz2, method="REML", correlation = corGaus(form = ~easting2+northing2, nugget = TRUE))
# 
# mod.sph = lme(log(mean_size) ~ sst_mean + fwd_mean + rugosity, random=~ 1|subregion/site, data=lapas_sz2, method="REML", correlation = corSpher(form = ~easting2+northing2, nugget = TRUE))
# 
# #mod.lin = lme(log(mean_size) ~ sst_mean + fwd_mean + rugosity, random=~ 1|subregion/site, data=lapas_sz2, method="REML", correlation = corLin(form = ~easting2+northing2, nugget = TRUE))
# ## mod.lin did not converge ##
# 
# mod.ratio = lme(log(mean_size) ~ sst_mean + fwd_mean + rugosity, random=~ 1|subregion/site, data=lapas_sz2, method="REML", correlation = corRatio(form = ~easting2+northing2, nugget = TRUE))
# 
# mod.exp = lme(log(mean_size) ~ sst_mean + fwd_mean + rugosity, random=~ 1|subregion/site, data=lapas_sz2, method="REML", correlation = corExp(form = ~easting2+northing2, nugget = TRUE))
# 
# cand.set.mod <- list(dens.g.REML, mod.gau, mod.sph,  mod.ratio, mod.exp)
# modnames.mod <-c("dens.g.REML","mod.gau", "mod.sph", "mod.ratio", "mod.exp")
# AICctab(cand.set.mod, mnames = modnames.mod, base=TRUE, weights=TRUE, logLik=TRUE)
# 
# lmtest::lrtest(dens.g.REML, mod.sph)

## Best model WITHOUT SPATIAL AUTOCORRELATION 
```

#### 
```{r gam}

## reducing number of samples to site level to decrease dispersion
lapas1 <- left_join(lapas_sz %>% select(-subregion) %>% aggregate(mean_size ~ site+region+easting+northing, ., mean),
                    lapas_sz %>% select(site, sst_mean, fwd_mean, rugosity) %>% distinct())

## check relationship shape
xyplot(lapas1$mean_size ~ lapas1$sst_mean, type=c("p", "smooth"), pch=21, cex=1.2, col="black", fill=c("light blue")) # sigmoid-like shape

xyplot(lapas1$mean_size ~ lapas1$fwd_mean, type=c("p", "smooth"), pch=21, cex=1.2, col="black", fill=c("light blue")) # flat line

xyplot(lapas1$mean_size ~ lapas1$rugosity, type=c("p", "smooth"), pch=21, cex=1.2, col="black", fill=c("light blue")) # sigmoid-like shape


# comparing LM, GAM with bs spline and SCAM with constrained shape (sigmoid)
lm_lp = lm(mean_size ~ sst_mean + fwd_mean + rugosity, data=lapas1)
gam_lp1 = mgcv::gam(mean_size ~ s(sst_mean) + s(fwd_mean) + s(rugosity), data=lapas1)
scam_lp = scam::scam(mean_size ~ s(sst_mean, bs="mpd") + s(fwd_mean) + s(rugosity, bs="mpd"), data=lapas1)

summary(lm_lp)
summary(gam_lp1)
summary(scam_lp)

AIC(lm_lp, gam_lp1, scam_lp)
# gam is better

gam_lp <- update(gam_lp1, .~.-s(fwd_mean)) # best model
gam_lp2 <- update(gam_lp, .~.-s(rugosity)+rugosity)
AIC(gam_lp, gam_lp2)

summary(gam_lp)
plot(gam_lp)


sim <- simulateResiduals(gam_lp) 
sim %>% plot() ## OK
sim %>% testDispersion() ## OK
sim %>% testUniformity() ## OK
sim %>% plotResiduals(lapas1$sst_mean, xlab = "SST mean", ylab = "scaled residuals") ## OK
sim %>% plotResiduals(lapas1$rugosity, xlab = "Rugosity", ylab = "scaled residuals") ## OK

## check spatial autocorrelation
sim %>% 
  testSpatialAutocorrelation(x = lapas1$easting, y = lapas1$northing, plot = T) 
# no spatial correlation

```

### Residuals
```{r}

par(mfrow = c(3,2))
mgcv::gam.check(gam_lp)

xyplot(resid(gam_lp, type="pearson") ~ sst_mean, data=lapas1, 
            ylab="Pearson residuals", xlab="SST", panel = function(x,y){
              panel.grid(h = -1, v = 2)
              panel.points(x, y, col = 1)
              panel.loess(x, y, span=0.5, col = 1, lwd=2)})

xyplot(resid(gam_lp, type="pearson") ~ rugosity, data=lapas1, 
            ylab="Pearson residuals", xlab="Roughness", panel = function(x,y){
              panel.grid(h = -1, v = 2)
              panel.points(x, y, col = 1)
              panel.loess(x, y, span=0.5, col = 1, lwd=2)})

```

#### Predictions
```{r predictions}
# sst_mean
lot.sst = ggeffects::ggpredict(gam_lp, c("sst_mean[all]"), type = c("fe"))
names(lot.sst) <- c("sst_mean", "predicted", "sd", "low", "high", "n")
 
sst_pred2 = ggplot(lapas1, aes(sst_mean, mean_size)) + geom_point(size= 2.2, alpha=0.3) + 
   geom_ribbon(data = lot.sst, aes(y = NULL, ymin = low, ymax = high), fill = "grey50", alpha=0.5)+
   geom_line(data = lot.sst, mapping = aes(x = sst_mean, y = predicted), size = 1.3, color="black")    +theme_classic() + theme(axis.text.x = element_text(size=12, color="black"), 
         axis.text.y = element_text(size=12, color="black"),
         axis.title.x = element_text(color="black", size=14),
         axis.title.y = element_text(color="black", size=16)) +
   labs(x = expression(paste("sea surface temperature (",degree,"C)")), y = "size (mm)") +
   theme(strip.text.x = element_text(size = 10), strip.text.y = element_text(size = 10))+
   theme(axis.ticks = element_line(colour = "black", size = 0.8)) +
   theme(axis.line = element_line(colour = 'black', size = 0.8)) +
   theme(axis.ticks.length = unit(4, "pt"), legend.position = "none")# 

 
# fwd_mean
lot.fwd = ggeffects::ggpredict(gam_lp, c("fwd_mean[all]"), type = c("fe"))
names(lot.fwd) <- c("fwd_mean", "predicted", "sd", "low", "high", "n")
 
fwd_pred2 = ggplot(lapas_sz, aes(fwd_mean, mean_size)) + 
   geom_point(size= 2.2, alpha=0.3) + 
   geom_ribbon(data = lot.fwd, aes(y = NULL, ymin = low, ymax = high), fill = "grey50", alpha=0.5)+
   geom_line(data = lot.fwd, mapping = aes(x = fwd_mean, y = predicted), size = 1.3, color="black")    +theme_classic() + theme(axis.text.x = element_text(size=12, color="black"), 
         axis.text.y = element_text(size=12, color="black"),
         axis.title.x = element_text(color="black", size=14),
         axis.title.y = element_text(color="black", size=16)) +
   labs(x = "freshwater discharge index", y = "size (mm)") +
   theme(strip.text.x = element_text(size = 10), strip.text.y = element_text(size = 10))+
   theme(axis.ticks = element_line(colour = "black", size = 0.8)) +
   theme(axis.line = element_line(colour = 'black', size = 0.8)) +
   theme(axis.ticks.length = unit(4, "pt"), legend.position = "none")
 
 
#rugosity
lot.rug = ggeffects::ggpredict(gam_lp, c("rugosity[all]"), type = c("fe"))
names(lot.rug) <- c("rugosity", "predicted", "sd", "low", "high", "n")
 
incl_pred2 = ggplot(lapas_sz, aes(rugosity, mean_size)) + 
   geom_point(size= 2.2, alpha=0.3) +#,aes(color=subregion)) + 
   geom_ribbon(data = lot.rug, aes(y = NULL, ymin = low, ymax = high), fill = "grey50", alpha=0.5)+
   geom_line(data = lot.rug, mapping = aes(x = rugosity, y = predicted), size = 1.3, color="black")    +theme_classic() + theme(axis.text.x = element_text(size=12, color="black"), 
         axis.text.y = element_text(size=12, color="black"),
         axis.title.x = element_text(color="black", size=14),
         axis.title.y = element_text(color="black", size=16)) +
   labs(x = "rugosity index", y = "size (mm)") +
   theme(strip.text.x = element_text(size = 10), strip.text.y = element_text(size = 10))+
   theme(axis.ticks = element_line(colour = "black", size = 0.8)) +
   theme(axis.line = element_line(colour = 'black', size = 0.8)) +
   theme(axis.ticks.length = unit(4, "pt"), legend.position = "none")

lapa_sz_preds = ggarrange(sst_pred2, fwd_pred2, incl_pred2, ncol = 2, nrow = 2, align = "hv")## arranging plots together ##
lapa_sz_preds
# ggsave("lapa_sz_preds.tiff", dpi = 600, compression = 'lzw')
```



#### Trying GAMM (but not sure about it)... maybe is better to continue with GLMM?
```{r GAMM}

# library(mgcv)
# library(DHARMa)
# 
# sz.gl.gam = gam(mean_size ~ s(sst_mean, bs='cs') + s(chl_mean, bs='cs') + s(wf_log, bs='cs') + s(ndvi, bs='cs') +s(fwd_mean, bs='cs') + s(rugosity, bs='cs') + s(incl_graus_tetra, bs='cs') +
#                     s(site, subregion, bs = 're'), data=lapas_sz, REML=T)
# summary(sz.gl.gam)
# gam.check(sz.gl.gam)
# 
# sz.gl.gam1 = update(sz.gl.gam, .~.-s(chl_mean, bs='cs'))
# summary(sz.gl.gam1)
# 
# sz.gl.gam2 = update(sz.gl.gam1, .~.-s(wf_log, bs='cs'))
# summary(sz.gl.gam2)
# 
# sz.gl.gam3 = update(sz.gl.gam2, .~.-s(ndvi, bs='cs'))
# summary(sz.gl.gam3)
# 
# sz.gl.gam4 = update(sz.gl.gam3, .~.-s(incl_graus_tetra, bs='cs'))
# summary(sz.gl.gam4)
# 
# sz.gl.gam5 = update(sz.gl.gam4, .~.-s(fwd_mean, bs='cs'))
# summary(sz.gl.gam5)
# 
# AIC(sz.gl.gam3, sz.gl.gam4, sz.gl.gam5)
# anova(sz.gl.gam4,sz.gl.gam5)
# 
# gam.check(sz.gl.gam5)
# plot.gam(sz.gl.gam5)
# 
# ## residuals / SOME IMPROOVEMENT BUT NOT ENOUGH
# SO.lp.sz.gam <- simulateResiduals(fittedModel = sz.gl.gam5)
# plot(SO.lp.sz.gam)
# 
# # ok
# plotResiduals(lapas_sz$sst_mean,  SO.lp.sz.gam$scaledResiduals)
# plotResiduals(lapas_sz$rugosity,  SO.lp.sz.gam$scaledResiduals)
# 
# # checking spatial dependency / NO SPATIAL DEPENDENCY IN GAM MODEL
# lapas_sz$easting2 = jitter(lapas_sz$easting) 
# lapas_sz$northing2 = jitter(lapas_sz$northing)
# testSpatialAutocorrelation(simulationOutput = SO.lp.sz.gam, x = lapas_sz$easting2, y= lapas_sz$northing2)
# 
# 
# # # # adding coordinates
# 
# lp.gam.sp = gam(mean_size ~ s(sst_mean, bs='cs') + s(rugosity, bs='cs') + te(easting, northing) + s(site, subregion, bs = 're'), data=lapas_sz, REML=T)
# 
# summary(lp.gam.sp)
# gam.check(lp.gam.sp)
# plot.gam(lp.gam.sp, scheme = 1)
# 
# ## residuals / BETTER WITH COORDINATES
# SO.lp.sz.gamSP <- simulateResiduals(fittedModel = lp.gam.sp)
# plot(SO.lp.sz.gamSP)
# 
# # ok
# plotResiduals(lapas_sz$sst_mean,  SO.lp.sz.gamSP$scaledResiduals)
# plotResiduals(lapas_sz$rugosity,  SO.lp.sz.gamSP$scaledResiduals)
# 
# testSpatialAutocorrelation(simulationOutput = SO.lp.sz.gamSP, x = lapas_sz$easting2, y= lapas_sz$northing2)

# CONCLUSION: SPATIAL CORRELATION WAS DETECTED FOR NON-SPATIAL GAM. SPATIAL GAM HAS BETTER RESIDUALS, BUT THERE IS STILL SOME PATTERNS
```



