---
title: "Nodilittorina.abundance_ALPS"
author: "ALPS"
date: "30 de março de 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Starting: loading packages and data
```{r message=FALSE, warning=FALSE, include=FALSE}

rm(list=ls())
set.seed(13)

# # # # # # # # # # # # # # # # # # # # # # # # 
# Models testing # ECHINOLITTORINA SIZE
# # # # # # # # # # # # # # # # # # # # # # # #

# load packages
library(glmmTMB)
library(tidyverse)
library(DHARMa)
library(ggeffects)
library(mgcv)
library(dplyr)
library(ggplot2)
library(ggpubr)

source("functions/HighstatLibV6.r")

# # # LOAD DATA
nodi_sz <- read.csv("data/nodi_sz2020.csv", header=T)

ordem <- unique(nodi_sz[order(nodi_sz$distance_S),]['site'])
nodi_sz$site <- factor(nodi_sz$site, levels = as.character(ordem$site))
nodi_sz$subregion <- factor(nodi_sz$subregion, levels = c("MRBS","SSCh","Ubatuba","SCRJ","MRRJ","LRRJ"))

rm(ordem)

summary(nodi_sz)

hist(nodi_sz$tamanho_mm)
sample(nodi_sz$tamanho_mm, 5000) %>% shapiro.test()
hist(log(sample(nodi_sz$tamanho_mm, 5000)))
shapiro.test(log(sample(nodi_sz$tamanho_mm, 5000)))
```

#### Plotting to check
```{r}
ggplot(nodi_sz, aes(tamanho_mm)) +
  geom_density(aes(fill=site), alpha=0.5) +
  #geom_histogram(aes(fill=site)) +
  facet_grid(subregion~., drop = TRUE) +
  theme_classic() +
  theme(legend.position = "")

ggplot(nodi_sz, aes(x = site, y = tamanho_mm, color = subregion)) + 
  geom_boxplot(notch=T) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

#### VIF and correlations
```{r}
nodi_sz %>% 
  na.omit() %>% 
  dplyr::select(-region, -subregion, -site, -easting, -northing, -ext, -distance_S, -incl_graus_chth, -ndvi) %>% 
  corvif()

nodi_sz %>% 
  dplyr::select(-region, -subregion, -site, -easting, -northing, -ext, -distance_S, -incl_graus_chth, -ndvi) %>% 
  PerformanceAnalytics::chart.Correlation(histogram=TRUE, pch=19)
```

#### Visual inspection
```{r}
nodi_sz %>% 
  dplyr::select(-region, -subregion, -easting, -northing, -ext, -distance_S, -incl_graus_chth, -ndvi, -ext_m_chth) %>% 
  group_by(site) %>% 
  summarise_if(is.numeric, mean, na.rm = TRUE) %>% 
  gather(key='fatores', value='valores', -site, -tamanho_mm) %>% 
  ggplot() + 
  geom_point(aes(x= valores, y=tamanho_mm)) + 
  geom_smooth(aes(x= valores, y=tamanho_mm)) + 
  facet_wrap(~ fatores, ncol=3, scales = 'free') +
  theme_classic()

nodi <- nodi_sz %>% 
  dplyr::select(-region, -subregion, -easting, -northing, -ext, -distance_S, -incl_graus_chth, -ndvi, -ext_m_chth) %>% 
  group_by(site) %>% 
  summarise_if(is.numeric, mean, na.rm = TRUE) %>% data.frame()

# offset
nodi_sz <- nodi_sz %>% 
  select(site, tamanho_mm) %>% 
  group_by(site) %>% 
  summarise(ofset = length(tamanho_mm)) %>% 
  full_join(nodi_sz, by='site') %>% 
  data.frame()
```


```{r}
size.n.full = glmmTMB(tamanho_mm ~ ndvi_sup_diff + sst_mean + chl_mean + fwd_mean + wf_log + inclinacao + rugosity + (1|region/subregion/site), data=nodi_sz, REML=T)

size.n.no_reg = glmmTMB(tamanho_mm ~ ndvi_sup_diff +  sst_mean + chl_mean + fwd_mean + wf_log + inclinacao + rugosity + (1|subregion/site), data=nodi_sz, REML=T)

size.n.no_sub = glmmTMB(tamanho_mm ~ ndvi_sup_diff +  sst_mean + chl_mean + fwd_mean + wf_log + inclinacao + rugosity + (1|region/site), data=nodi_sz, REML=T)

size.n.no_loc = glmmTMB(tamanho_mm ~ ndvi_sup_diff +  sst_mean + chl_mean + fwd_mean + wf_log + inclinacao + rugosity  + (1|region/subregion), data=nodi_sz, REML=T)

size.n.reg = glmmTMB(tamanho_mm ~ ndvi_sup_diff + sst_mean + chl_mean + fwd_mean + wf_log + inclinacao + rugosity  +(1|region), data=nodi_sz, REML=T)

size.n.sub = glmmTMB(tamanho_mm ~ ndvi_sup_diff + sst_mean + chl_mean + fwd_mean + wf_log + inclinacao + rugosity  +(1|subregion), data=nodi_sz, REML=T)

size.n.loc = glmmTMB(tamanho_mm ~ ndvi_sup_diff + sst_mean + chl_mean + fwd_mean + wf_log + inclinacao + rugosity  +(1|site), data=nodi_sz, REML=T)


cand.set.n <- list(size.n.full, size.n.no_reg, size.n.no_sub, size.n.no_loc, size.n.reg, size.n.sub, size.n.loc)
modnames.n<-c("size.n.full", "size.n.no_reg", "size.n.no_sub", "size.n.no_loc","size.n.reg", "size.n.sub", "size.n.loc")
AICctab(cand.set.n, mnames=modnames.n, base=TRUE, weights=TRUE, logLik=TRUE)## best model: size.n.no_reg and size.n.full

anova(size.n.no_reg, size.n.full) ## no significant difference, model dens.nb1.no_reg is simpler ##
## CONCLUSION: best random structure is subregion and site ##
```

#### Selecting fixed structure of predictive model (method = ML) 
```{r}
dens.n.ML = glmmTMB(tamanho_mm ~ ndvi_sup_diff + chl_mean + sst_mean + fwd_mean + wf_log + rugosity + inclinacao + (1|subregion/site), data=nodi_sz, REML=FALSE)
summary(dens.n.ML)
# nothing significant

# heteroscedasticity in residuals
library(lattice)
xyplot(resid(dens.n.ML, type="pearson") ~ fitted(dens.n.ML), 
       ylab="Pearson residuals", xlab="", panel = function(x,y){
         panel.grid(h = -1, v = 2)
         panel.points(x, y, col = 1)
         panel.loess(x, y, col = 1,lwd=2)})#

# testing boxcox transformation
boxcox_size <- caret::BoxCoxTrans(nodi_sz$tamanho_mm)
nodi_sz <- cbind(nodi_sz, size_BC=predict(boxcox_size, nodi_sz$tamanho_mm))

hist(nodi_sz$size_BC)## far better normal distribution ##
```

#### Selecting fixed structure of predictive model (method = ML) again (BC transformation)
```{r}
dens.n.MLBC = glmmTMB(size_BC ~ ndvi_sup_diff + chl_mean + sst_mean + fwd_mean + wf_log + rugosity + inclinacao + (1|subregion/site), data=nodi_sz, REML=FALSE)

summary(dens.n.MLBC) # nothing significant again

# improvement in residuals
xyplot(resid(dens.n.MLBC, type="pearson") ~ fitted(dens.n.MLBC), 
       ylab="Pearson residuals", xlab="", panel = function(x,y){
         panel.grid(h = -1, v = 2)
         panel.points(x, y, col = 1)
         panel.loess(x, y, col = 1,lwd=2)})#
```

#### Proceeding to model selection (fixed structure) (REML= F)
```{r}
# proceed model selection
dens.n.ML1 <- update(dens.n.MLBC, .~. -chl_mean)
summary(dens.n.ML1)

dens.n.ML2 <- update(dens.n.ML1, .~. -rugosity)
summary(dens.n.ML2)

dens.n.ML3 <- update(dens.n.ML2, .~. -inclinacao)
summary(dens.n.ML3)

dens.n.ML4 <- update(dens.n.ML3, .~. -fwd_mean) 
summary(dens.n.ML4)

dens.n.ML5 <- update(dens.n.ML4, .~. -ndvi_sup_diff)
summary(dens.n.ML5)

dens.n.ML6 <- update(dens.n.ML5, .~. -wf_log)
summary(dens.n.ML6)

dens.n.ML7 <- update(dens.n.ML6, .~. -sst_mean)
summary(dens.n.ML7) 

## CONCLUSION: none of predictive variables were selected ##
```

#### Refit with REML and summary 
```{r}
dens.n.REML = glmmTMB(size_BC ~ (1|subregion/site), data=nodi_sz, REML=TRUE)
## model doesn't converge ##

dens.n.REML2 = glmmTMB(size_BC ~ (1|region/subregion/site), data=nodi_sz, REML=TRUE)
## large portion of variance on region: put this factor back on model???

### What about testing model with random factors as fixed? ###

anova(dens.n.REML, dens.n.REML2) ## models are no different but model without REGION doesn't converge ##

sjPlot::plot_model(dens.n.REML2, type = "re", show.values = F, value.offset = .4)
```

#### Checking model assumptions 
```{r}
plot(dens.n.REML2) ## OK
qqnorm(resid(dens.n.REML2)); qqline(resid(dens.n.REML2)) ## OK
hist(resid(dens.n.REML2)) ## OK

xyplot(resid(dens.n.REML2, type="pearson") ~ fitted(dens.n.REML2), 
       ylab="Pearson residuals", xlab="Fitted values", panel = function(x,y){
         panel.grid(h = -1, v = 2)
         panel.points(x, y, col = 1)
         panel.loess(x, y, span=0.5, col = 1, lwd=2)})##

histogram(resid(dens.n.REML2, type="pearson"), col="white", xlab=list(label="Pearson residuals"))



```

#### Testing spatial structure on models
```{r}
# nodi_sz$easting2 = jitter(nodi_sz$easting) 
# nodi_sz$northing2 = jitter(nodi_sz$northing)
# 
# lmc <- lmeControl(niterEM = 5200, msMaxIter = 5200)
# 
# dens.n.REML1 = lme(size_BC ~ 1, random = ~1|subregion/site, data=nodi_sz, method = "REML", control= lmc)
# 
# dens.n.REML2b = lme(size_BC ~ 1, random = ~1|region/subregion/site, data=nodi_sz, method = "REML", control= lmc)
# 
# ### spatial models (considering 1|region/subregion/site or 1|subregion/site)
# env.gau = lme(size_BC ~ 1, random=~ 1|subregion/site, data=nodi_sz, method="REML",
#               correlation = corGaus(form = ~easting2+northing2, nugget = TRUE))
# 
# env.gaur = lme(size_BC ~ 1, random=~ 1|region/subregion/site, data=nodi_sz, method="REML",
#                correlation = corGaus(form = ~easting2+northing2, nugget = TRUE))
# 
# env.sph = lme(size_BC ~ 1, random=~ 1|region/subregion/site, data=nodi_sz, method="REML",
#               correlation = corSpher(form = ~easting2+northing2, nugget = TRUE))
# 
# env.sph.noR = lme(size_BC ~ 1, random=~ 1|subregion/site, data=nodi_sz, method="REML",
#                correlation = corSpher(form = ~easting2+northing2, nugget = TRUE))
# 
# env.lin = lme(size_BC ~ 1, random=~ 1|subregion/site, data=nodi_sz, method="REML",
#               correlation = corLin(form = ~easting2+northing2, nugget = TRUE))
# 
# env.linr = lme(size_BC ~ 1, random=~ 1|region/subregion/site, data=nodi_sz, method="REML",
#                correlation = corLin(form = ~easting2+northing2, nugget = TRUE), control=lmc)
# 
# env.ratio = lme(size_BC ~ 1, random=~ 1|subregion/site, data=nodi_sz, method="REML",
#                 correlation = corRatio(form = ~easting2 + northing2, nugget = TRUE), control= lmc)
# 
# env.ratior = lme(size_BC ~1, random=~ 1|region/subregion/site, data=nodi_sz, method="REML",
#                  correlation = corRatio(form = ~easting2+northing2, nugget = TRUE), control= lmc)
# 
# env.exp = lme(size_BC ~1, random=~ 1|subregion/site, data=nodi_sz, method="REML",
#               correlation = corExp(form = ~easting2+northing2, nugget = TRUE), control= lmc)
# 
# env.expr = lme(size_BC ~ 1, random=~ 1|region/subregion/site, data=nodi_sz, method="REML",
#                correlation = corExp(form = ~easting2+northing2, nugget = TRUE), control= lmc)
# 
# cand.set.env2 <- list(dens.n.REML1, dens.n.REML2, env.lin, env.gau, env.sph, env.ratio, env.exp, env.gaur, env.sph.noR, env.ratior, env.expr, env.linr) #env.lin
# modnames.env2<-c("dens.n.REML1", "dens.n.REML2", "env.lin", "env.gau", "env.sph", "env.ratio", "env.exp", "env.gaur", "env.sph.noR", "env.ratior", "env.expr", "env.linr") #"env.lin", 
# AICctab(cand.set.env2, mnames=modnames.env2, base=TRUE, weights=TRUE, logLik=TRUE)
# 
# anova(dens.n.REML1, dens.n.REML2b)
# ## no difference between random structure; best model dens.n.REML1##
# ## CONCLUSION: spatial structure did not improve models ## 
# ## What to do next??? ##
```

#### Plotting to check spatial patterns
```{r}
library(gstat)
library(sp)
RES   = residuals(dens.n.REML1, type="pearson")

mydata = data.frame(RES, nodi_sz$easting2, nodi_sz$northing2)

coordinates(mydata) <- c("nodi_sz.easting2","nodi_sz.northing2")

bubble(mydata, "RES", col = c("black","red"), alpha=0.1,
       main = "Pearson Residuals", xlab = "Northing",
       ylab = "Easting", scales = list(draw = T))  ## not clear

variogram1 <- variogram(RES ~ 1, mydata)
plot(variogram1) ## apparently there are some spatial patterns on residuals ##

### trying best spatial model (env.linr) ###
RES.lin   = residuals(env.linr, type="pearson")

mydata2 = data.frame(RES.lin, nodi_sz$easting2, nodi_sz$northing2)

coordinates(mydata2) <- c("nodi_sz.easting2","nodi_sz.northing2")

bubble(mydata2, "RES.lin", col = c("black","red"), alpha=0.1,
       main = "Pearson Residuals", xlab = "Northing",
       ylab = "Easting", scales = list(draw = T))  ## not clear

variogram2 <- variogram(RES.lin ~ 1, mydata2)
plot(variogram2) ## asame patterns as model without spatial correlation term ##
```


#### Trying GAM (not sure about it)
```{r}
nd.gl.gam.all = gam(size_BC ~ s(ndvi_sup_diff) + s(chl_mean) + s(sst_mean) + 
                      s(fwd_mean) + s(wf_log) + s(rugosity) + s(incl_graus_chth) +
                      s(site, subregion, bs = 're'), data=nodi_sz, REML=T)
summary(nd.gl.gam.all)

nd.gl.gam1 <- update(nd.gl.gam.all, .~.-s(ndvi_sup_diff))
summary(nd.gl.gam1)

nd.gl.gam2 <- update(nd.gl.gam1, .~.-s(rugosity))
summary(nd.gl.gam2)

nd.gl.gam3 <- update(nd.gl.gam2, .~.-s(chl_mean))
summary(nd.gl.gam3)

nd.gl.gam4 <- update(nd.gl.gam3, .~.-s(incl_graus_chth))
summary(nd.gl.gam4)

nd.gl.gam5 <- update(nd.gl.gam4, .~.-s(fwd_mean)+fwd_mean)
summary(nd.gl.gam5)

nd.gl.gam6 <- update(nd.gl.gam5, .~.-s(wf_log)+wf_log)
summary(nd.gl.gam6)

gam.check(nd.gl.gam6)
plot.gam(nd.gl.gam6, pages=1)

#
sim <- simulateResiduals(nd.gl.gam6) 
sim %>% plot() ## OK
sim %>% testDispersion() ## OK
sim %>% plotResiduals(nodi_sz$sst_mean, xlab = "SST mean", ylab = "scaled residuals") ## OK
sim %>% plotResiduals(nodi_sz$fwd_mean, xlab = "FWD mean", ylab = "scaled residuals") ## OK
sim %>% plotResiduals(nodi_sz$wf_log, xlab = "WF log", ylab = "scaled residuals") ## OK
sim %>% testSpatialAutocorrelation(x = nodi_sz$easting %>% jitter(), y = nodi_sz$northing %>% jitter(), plot = T) 

## spatial dependence detected and quantile deviations


## reduce to site level
abiot <- read.csv("~/Google Drive/PUBLICACOES/CONSUMERS_SE-BR/2021/data/abioticos_2020.csv", header = T)

nodi <- nodi_sz %>%
  group_by(site) %>% 
  #summarise(s_range = tamanho_mm - mean(tamanho_mm)) %>% 
  summarise(s_range = max(tamanho_mm) - min(tamanho_mm),
            tamanho_mm = mean(tamanho_mm),
            size_BC = mean(size_BC)) %>% 
  left_join(nodi_sz %>% select(region, subregion, easting, northing, site, distance_S, ndvi_sup_diff, chl_mean, sst_mean, fwd_mean, wf_log, rugosity, inclinacao) %>% distinct()) %>% 
  left_join(abiot[,c("site", "lat", "long")])


# GAM
m_gam <- gam(tamanho_mm ~ s(sst_mean) + s(ndvi_sup_diff) + s(fwd_mean) +
               s(chl_mean) + s(wf_log) + s(rugosity), data = nodi, REML = F)
summary(m_gam)

mod1 <- update(m_gam, .~.-s(ndvi_sup_diff))
summary(mod1)
mod2 <- update(mod1, .~.-s(rugosity))
summary(mod2)
mod3 <- update(mod2, .~.-s(chl_mean)+chl_mean)
summary(mod3)
mod4 <- gam(tamanho_mm ~ s(sst_mean, k=4) + s(fwd_mean) + s(wf_log, k=4), data = nodi, REML = F)
summary(mod4)

#
plot(mod3)
gam.check(mod4)
# maybe some covariance between variables

#
simulateResiduals(mod3) %>% plot() ## quantile deviance
simulateResiduals(mod3) %>% testDispersion() ## OK
simulateResiduals(mod3) %>% plotResiduals(nodi$sst_mean, xlab = "SST mean", ylab = "scaled residuals") ## OK
simulateResiduals(mod3) %>% plotResiduals(nodi$fwd_mean, xlab = "FWD mean", ylab = "scaled residuals") ## OK
simulateResiduals(mod3) %>% plotResiduals(nodi$wf_log, xlab = "WF log", ylab = "scaled residuals") ## OK
simulateResiduals(mod3) %>% testSpatialAutocorrelation(x = nodi$easting, y = nodi$northing, plot = T) 
# spatial dependence

## add coordinates
mod4 <- update(mod3, .~.+te(lat, long))
summary(mod4)
mod5 <- update(mod4, .~.-s(sst_mean))
summary(mod5)
mod6 <- update(mod5, .~.-chl_mean)
summary(mod6)
mod7 <- update(mod6, .~.-s(wf_log))
summary(mod7)
mod8 <- update(mod7, .~.-s(fwd_mean))
summary(mod8)

plot(mod8)
gam.check(mod8)

#
simulateResiduals(mod8) %>% plot() ## quantile deviance
simulateResiduals(mod8) %>% testDispersion() ## OK
simulateResiduals(mod8) %>% testSpatialAutocorrelation(x = nodi$easting, y = nodi$northing, plot = T) 
# no spatial dependence

plot(mod8)
plot(resid(mod8))
plot(resid(mod8), nodi$long, las=1)
plot(resid(mod8), nodi$lat, las=1)
qqnorm(resid(mod8)); qqline(resid(mod8)) ## OK
hist(resid(mod8)) ## not ok

```

```{r linear model}
# checking mean size per subregion, there is a clear difference between MRBS+SSCh and the rest
boxplot(tamanho_mm ~ subregion, data = nodi)

mod_lin0 <- lm(tamanho_mm ~ 1, data = nodi) 
summary(mod_lin)
mod_lin <- lm(tamanho_mm ~ lat + long, data = nodi)
# improvement in residuals with BC transformation
mod_lin1 <- lm(size_BC ~ lat + long, data = nodi)
summary(mod_lin1)
mod_lin2 <- lm(size_BC ~ lat, data = nodi)
summary(mod_lin2)

AIC(mod_lin1, mod_lin, mod_lin2)
anova(mod_lin1, mod_lin2)

plot(resid(mod_lin2))
plot(resid(mod_lin2), nodi$long, las=1)
plot(resid(mod_lin2), nodi$lat, las=1)
qqnorm(resid(mod_lin2)); qqline(resid(mod_lin2)) ## OK
hist(resid(mod_lin2)) ## OK
shapiro.test(resid(mod_lin2)) ## OK

#
nodi %>% 
  select(-region, -site, -size_BC:-region,- easting:-distance_S, -s_range) %>%
  aggregate(. ~ subregion, ., mean)

#### correlation
cor.test(nodi$tamanho_mm, nodi$lat)

ggplot(nodi, aes(lat, tamanho_mm)) + 
  geom_point(alpha=0.7, aes(color = subregion)) +
  theme_classic() +
  theme(legend.title = element_blank()) +
  geom_smooth(method = "lm") +
  scale_x_reverse() +
  labs(x = "Latitude (degrees)", y = "Size (cm)")

#####
# analysis of variance
anova <- lm(size_BC ~ ofset + region:subregion, nodi %>% left_join(nodi_sz %>% select(site, ofset) %>% distinct())) # %>% sjPlot::plot_models()
anova(anova)

model_means <- emmeans::emmeans(object = anova,
                       specs = ~ subregion | region) 

# add letters to each mean
model_means_cld <- multcomp::cld(object = model_means,
                       adjust = "Tukey",
                       Letters = letters,
                       alpha = 0.05) %>% 
  data.frame() %>% 
  filter(!is.na(emmean))
  

````

