---
title: "Str.siz_ALPS"
author: "ALPS"
date: "27-03-2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

#### Starting: loading packages and data
```{r message=FALSE, warning=FALSE, include=FALSE}

rm(list=ls())
set.seed(13)

library(tidyverse)
library(vegan)
library(mgcv)
library(lme4)
library(glmmTMB)
library(bbmle)
library(lmerTest)
library(lattice)
library(dplyr)
library(ggplot2)
library(nlme)
library(gridExtra)

source("functions/HighstatLibV6.r")## Codes for calculationg VIF ##

stramonita <- read.csv("data/stramonita_size.txt", header=T, sep="\t")

hist(stramonita$tamanho_mm)## distribution is fairly normal spite of shapiro-wilk test ##
shapiro.test(stramonita$tamanho_mm)
```

#### VIF and correlations
```{r}
stramonita %>% 
  na.omit() %>% 
  dplyr::select(-region, -subregion, -site, -easting, -northing, -distance_S) %>% 
  corvif()

stramonita %>% 
  dplyr::select(-site, -region, -subregion, -inclinacao, -northing, - easting, -distance_S, -Brach_RW) %>%
  PerformanceAnalytics::chart.Correlation(histogram=TRUE, pch=19)

# visual inspection
stramonita %>% 
  dplyr::select(-region, -subregion, -northing, - easting, -distance_S, - Brach_RW) %>% 
  gather(key='fatores', value='valores', -tamanho_mm, -site) %>% 
  ggplot() + 
  geom_point(aes(y= tamanho_mm, x=valores), alpha=0.5) + 
  geom_smooth(aes(y= tamanho_mm, x=valores)) +
  facet_wrap(~ fatores, ncol=3, scales = 'free') +
  theme_classic()
```

#### Selecting random structure of predictive model
```{r}
modS <- glmmTMB(tamanho_mm ~ chl_mean + sst_mean + fwd_mean + wf_log + rugosity + inclinacao+
               Tetr_size + Tetr_density + Brach_cover + Brach_size + perna_PA + 
               (1|site), data= stramonita, REML = T)# site

modSb <- glmmTMB(tamanho_mm ~ chl_mean + sst_mean + fwd_mean + wf_log + rugosity + inclinacao+
                Tetr_size + Tetr_density + Brach_cover + Brach_size + perna_PA + 
                (1|subregion), data= stramonita, REML = T)# subregion

modR <- glmmTMB(tamanho_mm ~ chl_mean + sst_mean + fwd_mean + wf_log + rugosity + inclinacao+
               Tetr_size + Tetr_density + Brach_cover + Brach_size + perna_PA + 
               (1|region), data= stramonita, REML = T)# region

modSSb <- glmmTMB(tamanho_mm ~ chl_mean + sst_mean + fwd_mean + wf_log + rugosity + inclinacao+
                 Tetr_size + Tetr_density + Brach_cover + Brach_size + perna_PA + 
                 (1|subregion/site), data= stramonita, REML = T)# subregion:site

modSR <- glmmTMB(tamanho_mm ~ chl_mean + sst_mean + fwd_mean + wf_log + rugosity + inclinacao+
                Tetr_size + Tetr_density + Brach_cover + Brach_size + perna_PA + 
                (1|region/site), data= stramonita, REML = T)# region:site

modSbR <- glmmTMB(tamanho_mm ~ chl_mean + sst_mean + fwd_mean + wf_log + rugosity + inclinacao+
                 Tetr_size + Tetr_density + Brach_cover + Brach_size + perna_PA + 
                 (1|region/subregion), data= stramonita, REML = T)# region:subregion

modSSbR <- glmmTMB(tamanho_mm ~ chl_mean + sst_mean + fwd_mean + wf_log + rugosity + inclinacao+
                  Tetr_size + Tetr_density + Brach_cover + Brach_size + perna_PA + 
                  (1|region/subregion/site), data= stramonita, REML = T)# region:subregion:site

cand.set.env <- list(modS, modSb, modR, modSSb, modSR, modSbR, modSSbR)
modnames.env<-c("modS", "modSb", "modR", "modSSb", "modSR", "modSbR", "modSSbR")
bbmle::AICctab(cand.set.env, mnames=modnames.env, base=TRUE, weights=TRUE, logLik=TRUE)
## CONCLUSION: Best model modS (only site as random term) ##
```

#### Selecting fixed structure of predictive model (method = ML) 
```{r}
mod1 <- glmmTMB(tamanho_mm ~ chl_mean + sst_mean + fwd_mean + wf_log + rugosity + inclinacao +
               Tetr_size + Tetr_density + Brach_cover + Brach_size + perna_PA + (1|site), data= stramonita, REML = F)
summary(mod1)

mod2 <- update(mod1, .~.-Tetr_size)
summary(mod2)

mod3 <- update(mod2, .~.-perna_PA)
summary(mod3)

mod4 <- update(mod3, .~.-Brach_size)
summary(mod4)

mod5 <- update(mod4, .~.-inclinacao)
summary(mod5)

mod5a <- update(mod5, .~.-rugosity)
summary(mod5a)

mod6 <- update(mod5a, .~.-Tetr_density)
summary(mod6)

mod7 <- update(mod6, .~.-sst_mean)
summary(mod7)

mod8 <- update(mod7, .~.-wf_log)
summary(mod8)

mod9 <- update(mod8, .~.-fwd_mean)
summary(mod9)

mod10 <- update(mod9, .~.-chl_mean)
summary(mod10)

## CONCLUSION m is the od10 model with all significant predictors (Brach_cover) ##
```

#### Refit with REML and summary 
```{r}
mod10.reml <- glmmTMB(tamanho_mm ~ Brach_cover + (1|site), data= stramonita, REML = TRUE) ## Final selected model ##
summary(mod10.reml)
```

#### Checking model assumptions 
```{r}
r1 = xyplot(resid(mod10.reml, type="pearson")~fitted(mod10.reml), 
       ylab="Pearson residuals", xlab="Fitted values", panel = function(x,y){
         panel.grid(h = -1, v = 2)
         panel.points(x, y, col = 1)
         panel.loess(x, y, span=0.5, col = 1, lwd=2)})## Plot residuals ~ fitted values with smoother##

r2 = xyplot(resid(mod10.reml, type="pearson")~Brach_cover, data=stramonita, 
       ylab="Pearson residuals", xlab="Mytilaster cover (%)", panel = function(x,y){
         panel.grid(h = -1, v = 2)
         panel.points(x, y, col = 1)
         panel.loess(x, y, span=0.5, col = 1, lwd=2)})## Plot residuals ~ fitted values with smoother##

r3 = histogram(resid(mod10.reml, type="pearson"), col="white", xlab=list(label="Pearson residuals"))

all.res = grid.arrange(r1, r2, r3, ncol=1) ### plotting together ###
## CONCLUSION: there are no problems on residuals ##
```

#### Plotting to check spatial autocorrelation
```{r}
stramonita2= as.data.frame(stramonita)
##create jitter (values a little different) because correlation term doesn't work when distance is 0 between replicates###
stramonita2$easting2 = jitter(stramonita2$easting)
stramonita2$northing2 = jitter(stramonita2$northing)

library(gstat)
library(sp)
library(ncf)

## getting the residuals ##
E <- residuals(mod10.reml, type="pearson") 
## creating dataframe with residuals and coordinates ##
mydata <- data.frame(E, stramonita$easting, stramonita$northing) 
## creating spatial object ##
coordinates(mydata) <- c("stramonita.easting","stramonita.northing") 
## plotting bubblemap with residuals ##
bubble(mydata, "E", col = c("black","grey"), alpha=0.4,
       main = "Pearson Residuals", xlab = "Northing",
       ylab = "Easting", scales = list(draw = T)) 

vario1 <- variogram(E ~ 1, mydata)
plot(vario1)

correlog <- spline.correlog(x = stramonita[, "easting"],
                            y = stramonita[, "northing"],
                            z = residuals(mod10.reml, type = "pearson"))
plot(correlog) ### NO SPATIAL AUTOCORRELATION ### NOT sure about this function ###

#
coords <- stramonita %>% 
  select(site, easting, northing) %>% 
  distinct()

simulateResiduals(mod10.reml) %>% 
  recalculateResiduals(group = stramonita$site) %>% 
  testSpatialAutocorrelation(x = coords$easting, y = coords$northing, plot = T)

## CONCLUSION: apparently there are not spatial patterns on residuals ##
```

#### Testing for spatial autocorrelation 
```{r}
# mod10.reml <- lmer(tamanho_mm ~ Brach_cover + (1|site), data= stramonita, REML = TRUE)
# 
# mod10.gau = lme(tamanho_mm ~ Brach_cover, random=~ 1|site, data=stramonita2, method="REML",
#               correlation = corGaus(form = ~easting2+northing2, nugget = TRUE))
# 
# mod10.sph = lme(tamanho_mm ~ Brach_cover, random=~ 1|site, data=stramonita2, method="REML",
#               correlation = corSpher(form = ~easting2+northing2, nugget = TRUE))
# 
# mod10.lin = lme(tamanho_mm ~ Brach_cover, random=~ 1|site, data=stramonita2, method="REML",
#               correlation = corLin(form = ~easting2+northing2, nugget = TRUE))
# 
# mod10.ratio = lme(tamanho_mm ~ Brach_cover, random=~ 1|site, data=stramonita2, method="REML",
#                 correlation = corRatio(form = ~easting2+northing2, nugget = TRUE))
# 
# mod10.exp = lme(tamanho_mm ~ Brach_cover, random=~ 1|site, data=stramonita2, method="REML",
#               correlation = corExp(form = ~easting2+northing2, nugget = TRUE))
# 
# cand.set.mod10 <- list(mod10.reml, mod10.gau, mod10.sph, mod10.lin, mod10.ratio, mod10.exp)
# modnames.mod10 <-c("mod10.reml","mod10.gau", "mod10.sph", "mod10.lin", "mod10.ratio", "mod10.exp")
# AICctab(cand.set.mod10, mnames = modnames.mod10, base=TRUE, weights=TRUE, logLik=TRUE)
## Best model WITHOUT SPATIAL AUTOCORRELATION 
```

#### Predictions
```{r}
library(ggeffects)
library(ggplot2)
library(ggpubr)

plot_model(mod10.reml, type = "pred")

#
str_brac = ggeffects::ggpredict(mod10.reml, c("Brach_cover[all]"), type = c("fe"))
names(str_brac) <- c("Brach_cover", "predicted", "sd", "low", "high", "n")

str_pred = ggplot(stramonita, aes(Brach_cover, tamanho_mm)) + 
           geom_point(size= 2.2, alpha=0.3) + 
           geom_ribbon(data = str_brac, aes(y = NULL, ymin = low, ymax = high), fill = "grey50", alpha=0.5)+
           geom_line(data = str_brac, mapping = aes(x = Brach_cover, y = predicted), size = 1.3, color="black") +
           theme_classic() + theme(axis.text.x = element_text(size=12, color="black"), 
           axis.text.y = element_text(size=12, color="black"),
           axis.title.x = element_text(color="black", size=14),
           axis.title.y = element_text(color="black", size=16)) +
      labs(x = expression(paste(italic("M. solisianus")," cover (%)")), y = ("size (mm)")) +
      theme(strip.text.x = element_text(size = 10), strip.text.y = element_text(size = 10))+
      scale_y_continuous(breaks = seq(0, 60, 15)) + scale_x_continuous(breaks = seq(0, 100, 25))+
      theme(axis.ticks = element_line(colour = "black", size = 0.8)) + expand_limits(x=c(0,100), y=c(0, 60))+
      theme(axis.line = element_line(colour = 'black', size = 0.8)) +
      theme(axis.ticks.length = unit(4, "pt"), legend.position = "")

str_pred

ggsave("str_pred.tiff", dpi = 600, compression = 'lzw')
```

#### Trying GAMM (but not sure about it)... maybe is better to continue with GLMM?
```{r}
# # fixed structure

# # all variables
# mod.gam <- gam(tamanho_mm ~ s(site, bs="re") + s(chl_mean) + s(sst_mean) + s(fwd_mean) + s(wf_log) + s(rugosity) + 
#                  s(Tetr_size) + s(Tetr_density) + s(Brach_cover) + s(Brach_size) + perna_PA, 
#                data= stramonita, na.action = na.exclude, method = 'REML')
# summary(mod.gam)
# 
# # remove splines from linear fits
# mod.gam1 <- gam(tamanho_mm ~ s(site, bs="re") + s(chl_mean, bs='cs') + s(sst_mean, bs='cs') + s(fwd_mean, bs='cs') + s(wf_log, bs='cs') + s(rugosity, bs='cs') + 
#                   s(Tetr_size, bs='cs') + s(Tetr_density, bs='cs') + s(Brach_cover, bs='cs') + s(Brach_size, bs='cs') + perna_PA, 
#                 data= stramonita, na.action = na.exclude, method = 'REML')
# summary(mod.gam1)
# 
# # variable selection
# mod1g <- update(mod.gam1, .~.-s(Brach_size, bs='cs'))
# summary(mod1g)
# mod2g <- update(mod1g, .~.-s(Tetr_size, bs='cs'))
# summary(mod2g)
# mod3g <- update(mod2g, .~.-s(Tetr_density, bs='cs'))
# summary(mod3g)
# mod4g <- update(mod3g, .~.-s(fwd_mean, bs='cs'))
# summary(mod4g)
# mod5g <- update(mod4g, .~.-s(rugosity, bs='cs'))
# summary(mod5)
# mod6g <- update(mod5g, .~.-s(chl_mean, bs='cs'))
# summary(mod6g)
# mod7g <- update(mod6g, .~.-perna_PA)
# summary(mod7g)
# mod8g <- update(mod7g, .~.-s(sst_mean, bs='cs'))
# summary(mod8g)
# mod9g <- update(mod8g, .~.-s(Brach_cover, bs='cs')+Brach_cover)
# summary(mod9g)
# 
# plot.gam(mod9g, all.terms = T, pages=1)
# gam.check(mod9g)
# anova(mod9g)
# 
# AIC(mod8g, mod9g)
# AIC(mod10.reml, mod9g)
# 
# # residuals
# xyplot(resid(mod9g, type="pearson")~fitted(mod9g), 
#        ylab="Pearson residuals", xlab="Fitted values", panel = function(x,y){
#          panel.grid(h = -1, v = 2)
#          panel.points(x, y, col = 1) #span = 0.2,
#          panel.loess(x, y, col = 1,lwd=2)})## 
# 
# histogram(resid(mod9g, type="pearson"), col="white", xlab=list(label="Pearson residuals"))
# 
# 
# ##### prediction
# # wf_log
# str.wfg = ggeffects::ggpredict(mod9g, c("wf_log[all]"), type = c("fe"))
# names(str.wfg) <- c("wf_log", "predicted", "sd", "low", "high", "n")
# 
# ggplot(stramonita, aes(wf_log, tamanho_mm)) + 
#   geom_point(size= 2.2, alpha=0.3) + 
#   #geom_point(size= 2.2, alpha=0.8, aes(color=subregion)) + 
#   geom_ribbon(data = str.wfg, aes(y = NULL, ymin = low, ymax = high), fill = "grey50", alpha=0.5)+
#   geom_line(data = str.wfg, mapping = aes(x = wf_log, y = predicted), size = 1.3, color="black") +
#   theme_classic() +
#   theme(axis.text.x = element_text(size=12, color="black"), 
#         axis.text.y = element_text(size=12, color="black"),
#         axis.title.x = element_text(color="black", size=14),
#         axis.title.y = element_text(color="black", size=16)) +
#   #labs(x = expression(paste("Chlorophyll-a (",mg.m^-3,")")), y = expression(paste("Density ", "(", ind.100cm^-2, ")"))) +
#   theme(strip.text.x = element_text(size = 10), strip.text.y = element_text(size = 10))+
#   #scale_y_continuous(breaks = seq(0, 16, 2.0)) +
#   theme(axis.ticks = element_line(colour = "black", size = 0.8)) +
#   theme(axis.line = element_line(colour = 'black', size = 0.8)) +
#   theme(axis.ticks.length = unit(4, "pt"), legend.position = "")
# 
# # brach cover
# str.bcg = ggeffects::ggpredict(mod9g, c("Brach_cover[all]"), type = c("fe"))
# names(str.bcg) <- c("Brach_cover", "predicted", "sd", "low", "high", "n")
# 
# ggplot(stramonita, aes(Brach_cover, tamanho_mm)) + 
#   geom_point(size= 2.2, alpha=0.3) + 
#   #geom_point(size= 2.2, alpha=0.8, aes(color=subregion)) + 
#   geom_ribbon(data = str.bcg, aes(y = NULL, ymin = low, ymax = high), fill = "grey50", alpha=0.5)+
#   geom_line(data = str.bcg, mapping = aes(x = Brach_cover, y = predicted), size = 1.3, color="black") +
#   theme_classic() +
#   theme(axis.text.x = element_text(size=12, color="black"), 
#         axis.text.y = element_text(size=12, color="black"),
#         axis.title.x = element_text(color="black", size=14),
#         axis.title.y = element_text(color="black", size=16)) +
#   #labs(x = expression(paste("Chlorophyll-a (",mg.m^-3,")")), y = expression(paste("Density ", "(", ind.100cm^-2, ")"))) +
#   theme(strip.text.x = element_text(size = 10), strip.text.y = element_text(size = 10))+
#   #scale_y_continuous(breaks = seq(0, 16, 2.0)) +
#   theme(axis.ticks = element_line(colour = "black", size = 0.8)) +
#   theme(axis.line = element_line(colour = 'black', size = 0.8)) +
#   theme(axis.ticks.length = unit(4, "pt"), legend.position = "")
# 
# 
# ##################
# # # add coordinates
# mod.str.sp <- gam(tamanho_mm ~ s(Brach_cover, bs="cs") + s(wf_log, bs='cs') + s(easting, northing), 
#                 data= stramonita, na.action = na.exclude, method = 'REML')
# mod.str.spt <- gam(tamanho_mm ~ s(Brach_cover, k=5) + s(easting, northing) + ti(easting, northing, Brach_cover), 
#                   data= stramonita, na.action = na.exclude, method = 'REML')
# 
# summary(mod.str.sp)
# plot(mod.str.sp, pages=1)
# gam.check(mod.str.sp)
# anova(mod.str.sp)
# plot(resid(mod.str.sp), las=1)
# 
# mod.str.sp1 <- gam(tamanho_mm ~s(wf_log, k=3) + s(easting, northing), #select = T,
#                   data= stramonita, na.action = na.exclude, method = 'REML')
# mod.str.spt1 <- gam(tamanho_mm ~ s(wf_log, bs='gp') + s(easting, northing) + ti(easting, northing, wf_log), 
#                    data= stramonita, na.action = na.exclude, method = 'REML')
# 
# summary(mod.str.spt1)
# plot(mod.str.spt1, pages=1, residuals = T)
# gam.check(mod.str.spt1)
# anova(mod.str.sp1)
# gam.vcomp(mod.str.spt1)
# 
# 
# # CONCLUSION LINEAR MODEL IS MORE ROBUST AND EASIER TO EXPLAIN
# AIC(mod10, mod.lme.lm, mod.str.sp, mod.str.spt1, mod.str.spt)
```
