---
title: "Tetraclita.size"
author: "Cesar Cordeiro"
date: "3/31/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE, paged.print=TRUE}

rm(list=ls())
set.seed(13)

# load packages
library(glmmTMB)
library(tidyverse)
library(DHARMa)
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)

# source("script/HighstatLibV6.r")


tetra_sz1 <- read.csv("data/tetra_sz2020.csv", header=T)

tetra_sz <- tetra_sz1 %>% 
  group_by(site, quadrado) %>% 
  summarise(mean_size = mean(tamanho_mm)) %>% 
  dplyr::select(site, mean_size) %>% 
  inner_join(., unique(tetra_sz1[,c(1,4:17)]), by='site') %>% 
  data.frame()

ordem <- unique(tetra_sz[order(tetra_sz$distance_S),]['site'])
tetra_sz$site <- factor(tetra_sz$site, levels = as.character(ordem$site))
tetra_sz$subregion <- factor(tetra_sz$subregion, levels = c("MRBS","SSCh","Ubatuba","SCRJ","MRRJ","LRRJ"))

rm(tetra_sz1, ordem)

# removing outliers improve model fitting
tetra_sz <- tetra_sz %>% 
  filter(mean_size < mean(mean_size) + sd(mean_size) & mean_size > mean(mean_size) - sd(mean_size))

```

#### Plotting for exploration
```{r data exploration}

hist(tetra_sz$mean_size)
shapiro.test(tetra_sz$mean_size)
hist(log(tetra_sz$mean_size))
shapiro.test(log(tetra_sz$mean_size))


ggplot(tetra_sz, aes(mean_size)) +
  geom_density(aes(fill=site), alpha=0.5) +
  facet_grid(subregion~., drop = TRUE) +
  theme_classic() +
  theme(legend.position = "")

ggplot(tetra_sz, aes(x = site, y = mean_size, color = subregion)) + 
  geom_boxplot(notch=T) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x="", "Size (mm)")

```

#### VIF and correlations
```{r VIF}

tetra_sz %>% 
  na.omit() %>% 
  dplyr::select(-region,-subregion,-site,-easting,-northing, -distance_S, -incl_graus_tetra) %>% 
  corvif()

tetra_sz %>% 
  dplyr::select(-region,-subregion,-site, -easting, -northing, -distance_S, -incl_graus_tetra) %>% 
  PerformanceAnalytics::chart.Correlation(histogram=TRUE, pch=19)

# visual inspection
tetra_sz %>% 
  dplyr::select(-region,-subregion, -easting, -northing, -distance_S) %>% 
  group_by(site) %>% 
  summarise_if(is.numeric, mean, na.rm = TRUE) %>% 
  gather(key='fatores', value='valores', -site, -mean_size) %>% 
  ggplot() + 
  geom_point(aes(x= valores, y=mean_size)) + 
  geom_smooth(aes(x= valores, y=mean_size)) + 
  facet_wrap(~ fatores, ncol=3, scales = 'free') +
  theme_classic()

```



#### checking log-transformation

```{r log-trasformation}
full.l = glmmTMB(log(mean_size) ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab + (1|subregion/site), data=tetra_sz, REML=T) # singular fit

full = glmmTMB(mean_size ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab + (1|subregion/site), data=tetra_sz, REML=T)

full.l %>% resid() %>% hist
full.l %>% resid() %>% plot

full %>% resid() %>% hist
full %>% resid() %>% plot

# Models with (1|region/subregion/site) did not converge
# Better log-transform

```

#### Selecting random structure of predictive model

```{r random, Distribution, message=FALSE, warning=FALSE}

full = glmmTMB(log(mean_size) ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab + (1|region/subregion/site), data=tetra_sz, REML=T)
no_reg = glmmTMB(log(mean_size) ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab +(1|subregion/site), data=tetra_sz, REML=T)
no_sub = glmmTMB(log(mean_size) ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab +(1|region/site), data=tetra_sz, REML=T)
no_loc = glmmTMB(log(mean_size) ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab +(1|region/subregion), data=tetra_sz, REML=T)
reg = glmmTMB(log(mean_size) ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab +(1|region), data=tetra_sz, REML=T)
sub = glmmTMB(log(mean_size) ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab +(1|subregion), data=tetra_sz, REML=T)
loc = glmmTMB(log(mean_size) ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab +(1|site), data=tetra_sz, REML=T)

cand.set.env <- list(full, no_reg, no_sub, no_loc, reg, sub, loc)
modnames.env<-c("full","no_reg","no_sub", "no_loc","reg", "sub", "loc")
AICctab(cand.set.env, mnames=modnames.env, base=TRUE, weights=TRUE, logLik=TRUE)

anova(no_sub, loc)

## CONCLUSION: Best random structure 'loc' (site only) ##
```

#### Selecting fixed structure of predictive model (method = ML)  
```{r fixed, message=FALSE, warning=FALSE}

size.t.ML = glmmTMB(log(mean_size) ~ str_ab + chl_mean + sst_mean + fwd_mean + wf_log + rugosity + incl_graus_tetra +
                     (1|site), data=tetra_sz, REML=F)
summary(size.t.ML)

size.t.ML1 <- update(size.t.ML, .~. -chl_mean)
summary(size.t.ML1)

size.t.ML2 <- update(size.t.ML1, .~. -str_ab)
summary(size.t.ML2) 

size.t.ML3 <- update(size.t.ML2, .~. -wf_log)
summary(size.t.ML3) 

size.t.ML4 <- update(size.t.ML3, .~. -rugosity)
summary(size.t.ML4) 

size.t.ML5 <- update(size.t.ML4, .~. -incl_graus_tetra)
summary(size.t.ML5) 

size.t.ML6 <- update(size.t.ML5, .~. -sst_mean)
summary(size.t.ML6) 

anova(size.t.ML5, size.t.ML6)
## CONCLUSION: Best model includes only FWD (size.t.ML6) ##
```

#### Refit with REML and summary 
```{r REML}

size.t.REML = glmmTMB(log(mean_size) ~ fwd_mean + (1|site), data=tetra_sz, REML=T)
summary(size.t.REML)

# random effects
sjPlot::plot_model(size.t.REML, type = "re") # random effects
sjPlot::plot_model(size.t.REML, type = "pred") # random effects


```

#### Checking model assumptions 
```{r residuals}

tetra.s.res <- xyplot(resid(size.t.REML, type="pearson") ~ fitted(size.t.REML), 
               ylab="Pearson residuals", xlab="Fitted values", 
               panel = function(x,y){
                 panel.grid(h = -1, v = 2)
                 panel.points(x, y, col = 1)
                 panel.loess(x, y, span = 0.5, col = 1,lwd=2)}) ## Plot residuals ~ fitted values with smoother##

tetra.s.fwd <- xyplot(resid(size.t.REML, type="pearson") ~ fwd_mean, data=tetra_sz, 
                    ylab="Pearson residuals", xlab="Freshwater influence index",
                    panel = function(x,y){
                      panel.grid(h = -1, v = 2)
                      panel.points(x, y, col = 1)
                      panel.loess(x, y, span = 0.5, col = 1,lwd=2)})

hist_res <- histogram(resid(size.t.REML, type="pearson"), col="white", xlab=list(label="Pearson residuals"))

grid.arrange(tetra.s.res, tetra.s.fwd, hist_res, ncol=3)

### checking residuals against all variables
tetra_sz %>% 
  dplyr::select(site, chl_mean, sst_mean, fwd_mean, wf_log, rugosity, inclinacao, str_ab, str_size, mean_size) %>% 
  pivot_longer(names_to = 'fatores', values_to ='valores', cols=chl_mean:str_size) %>% 
  #aggregate(.~site+fatores, FUN=mean, data=.) %>% 
  bind_cols(resid = rep(resid(size.t.REML,  type="pearson"), 8)) %>% 
  aggregate(. ~ site + fatores, FUN=mean, data=.) %>% 
  ggplot() + 
  geom_point(aes(x= valores, y=resid)) + 
  geom_smooth(aes(x= valores, y=resid)) +
  facet_wrap(~ fatores, ncol=3, scales = 'free') +
  theme_classic()


##
simulateResiduals(size.t.REML) %>% plot() ## Deviation, outlier and upper quantile
simulateResiduals(size.t.REML) %>% testDispersion() ## OK
simulateResiduals(size.t.REML) %>% plotResiduals(tetra_sz$fwd_mean, xlab = "FWD mean", ylab = "scaled residuals") ## some deviance in upper quantile

```

#### Plotting to check spatial autocorrelation
```{r check spatial}

library(gstat)
library(sp)
library(ncf)

##create jitter (values a little different) because correlation term doesn't work when distance is 0 between replicates###
tetra_sz$easting2 =  jitter(tetra_sz$easting)
tetra_sz$northing2 = jitter(tetra_sz$northing)

## getting the residuals ##
E <- residuals(size.t.REML, type="pearson") 
## creating dataframe with residuals and coordinates ##
mydata <- data.frame(E, tetra_sz$easting, tetra_sz$northing) 
## creating spatial object ##
coordinates(mydata) <- c("tetra_sz.easting","tetra_sz.northing") 
## plotting bubblemap with residuals ##
bubble(mydata, "E", col = c("black","red"), alpha=0.1,
       main = "Pearson Residuals", xlab = "Northing",
       ylab = "Easting", scales = list(draw = T)) 

vario1 <- variogram(E ~ 1, mydata)
plot(vario1) #, ylim = c(0.25,0.32)
## CONCLUSION: Some spatial autocorrelation ##

# recalculate residuals per site
coords <- tetra_sz %>% 
  select(site, easting, northing) %>% 
  distinct()
  
simulateResiduals(size.t.REML) %>% 
  recalculateResiduals(group = tetra_sz$site) %>% 
  testSpatialAutocorrelation(x = coords$easting, y = coords$northing, plot = T)

## no spatial correlation at the site level

```


#### Fitting spatial models (need to check)
```{r spatial model}

# env.REML = lme(log(mean_size) ~ fwd_mean, random = ~1|site, data = tetra_sz, method="REML")
# 
# env.gau = lme(log(mean_size) ~  fwd_mean, random=~ 1|site, data=tetra_sz, method="REML",
#               correlation = corGaus(form = ~easting2+northing2, nugget = TRUE))
# 
# env.sph = lme(log(mean_size) ~ fwd_mean, random=~ 1|site, data=tetra_sz, method="REML",
#               correlation = corSpher(form = ~easting2+northing2, nugget = TRUE))
# 
# env.lin = lme(log(mean_size) ~ fwd_mean, random=~ 1|site, data=tetra_sz, method="REML", 
#               correlation = corLin(form = ~easting2+northing2, nugget = TRUE)) 
# 
# env.ratio = lme(log(mean_size) ~ fwd_mean, random=~ 1|site, data=tetra_sz, method="REML",
#                 correlation = corRatio(form = ~easting2+northing2, nugget = TRUE))
# 
# env.exp = lme(log(mean_size) ~ fwd_mean, random=~ 1|site, data=tetra_sz, method="REML",
#               correlation = corExp(form = ~easting2+northing2, nugget = TRUE))
# 
# cand.set.env2 <- list(env.REML, env.gau, env.sph, env.lin, env.ratio, env.exp)
# modnames.env2<-c("env.REML","env.gau","env.sph","env.lin","env.ratio", "env.exp")
# AICctab(cand.set.env2, mnames=modnames.env2, base=TRUE, weights=TRUE, logLik=TRUE)
# 
# anova(env.gau, env.REML)

# Spatial structure did not enhance model
```
#### Simplifying model

```{r linear}
## reducing number of samples to site level to decrease dispersion
tetra1 <- left_join(tetra_sz %>% select(-str_ab, -subregion) %>% aggregate(mean_size ~ site+region+easting+northing, ., mean),
                    tetra_sz %>% select(-str_ab, -subregion) %>% aggregate(fwd_mean ~ site+region+easting+northing, ., mean))

## check relationship shape
xyplot(tetra1$mean_size ~ tetra1$fwd_mean,type=c("p", "smooth"), pch=21, cex=1.2, col="black", fill=c("light blue")) # sigmoid-like shape

# comparing GAM with bs spline and SCAM with constrained shape (sigmoid)
lm_lp = lm(mean_size ~ fwd_mean, data=tetra1)
gam_lp = mgcv::gam(mean_size ~ s(fwd_mean), data=tetra1) # overfitting, lowering K turns to linear
scam_lp = scam::scam(mean_size ~ s(fwd_mean, bs="mpd"), data=tetra1)

summary(lm_lp)
summary(gam_lp)
summary(scam_lp)
AIC(lm_lp, gam_lp, scam_lp)
# final models are linear

# check assumptions
simulateResiduals(lm_lp) %>% plot() ## OK
simulateResiduals(lm_lp) %>% testDispersion() ## OK
simulateResiduals(lm_lp) %>% testUniformity() ## OK
simulateResiduals(lm_lp) %>% plotResiduals(tetra1$fwd_mean, xlab = "FWD mean", ylab = "scaled residuals") ## small deviance in lower and upper quantiles

## check spatial autocorrelation
simulateResiduals(lm_lp) %>% 
  testSpatialAutocorrelation(x = tetra1$easting, y = tetra1$northing, plot = T) 
# no autocorrelation

```

#### Checking model assumptions 
```{r}
r1 = xyplot(resid(lm_lp, type="pearson") ~ fitted(lm_lp), 
       ylab="Pearson residuals", xlab="Fitted values", panel = function(x,y){
         panel.grid(h = -1, v = 2)
         panel.points(x, y, col = 1)
         panel.loess(x, y, span=0.5, col = 1, lwd=2)})## Plot residuals ~ fitted values with smoother##

r2 = xyplot(resid(lm_lp, type="pearson") ~ fwd_mean, data=tetra1, 
       ylab="Pearson residuals", xlab="log(Wave fetch)", panel = function(x,y){
         panel.grid(h = -1, v = 2)
         panel.points(x, y, col = 1)
         panel.loess(x, y, span=0.5, col = 1, lwd=2)})## Plot residuals ~ fitted values with smoother##

r3 = histogram(resid(lm_lp, type="pearson"), col="white", xlab=list(label="Pearson residuals"))

grid.arrange(r1, r2, r3, ncol=2) ### plotting together ###


```
#### Predictions of best model
```{r predictions}

library(ggeffects)
library(ggplot2)
library(ggpubr)

# fwd_mean
sjPlot::plot_model(lm_lp, type = "pred")

# 
# env.fwd = ggpredict(lm_lp, c("fwd_mean[all]"), type = c("fe"))
# names(env.fwd) <- c("fwd_mean", "predicted", "sd", "low", "high", "n")
# 
# fwd.plot = ggplot(tetra_sz, aes(fwd_mean, mean_size)) + 
#   geom_point(size= 2.2, alpha=0.3) + #alpha=0.8, aes(color=subregion) 
#   geom_ribbon(data = env.fwd, aes(y = NULL, ymin = low, ymax = high), fill = "grey50", alpha=0.5)+
#   geom_line(data = env.fwd, mapping = aes(x = fwd_mean, y = predicted), size = 1.3, color="black") +
#   theme_classic() + 
#   theme(axis.text.x = element_text(size=12, color="black"), 
#         axis.text.y = element_text(size=12, color="black"),
#         axis.title.x = element_text(color="black", size=14),
#         axis.title.y = element_text(color="black", size=16)) +
#   labs(x = "Freshwater discharge index", y = "Tetraclita size  (mm)") +
#   theme(strip.text.x = element_text(size = 10), strip.text.y = element_text(size = 10))+
#   scale_y_continuous(breaks = seq(0, 9, 1.0)) +
#   theme(axis.ticks = element_line(colour = "black", size = 0.8)) +
#   theme(axis.line = element_line(colour = 'black', size = 0.8)) +
#   theme(axis.ticks.length = unit(4, "pt"))
# 
# fwd.plot

ggsave("Tetr_s_preds.tiff", dpi = 600, compression = 'lzw')
```
