---
title: "Tetraclita.density"
author: "Cesar Cordeiro"
date: "3/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

#### Starting: loading packages and data
```{r message=FALSE, warning=FALSE, paged.print=TRUE}

rm(list=ls())
set.seed(13)

# load packages
library(lme4)
library(bbmle)
library(glmmTMB)
library(tidyverse)
library(nlme)
library(DHARMa)
library(lmerTest)
library(nlme)
library(lattice)
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)

# source("script/HighstatLibV6.r")


tetra_ab <- read.csv("data/tetra_ab2020.csv", header=T)
summary(tetra_ab)

ordem <- unique(tetra_ab[order(tetra_ab$distance_S),]['site'])
tetra_ab$site <- factor(tetra_ab$site, levels = as.character(ordem$site))
tetra_ab$subregion <- factor(tetra_ab$subregion, levels = c("MRBS","SSCh","Ubatuba","SCRJ","MRRJ","LRRJ"))
tetra_ab$str_ab[is.na(tetra_ab$str_ab)] <- 0.01

tetra_ab <- tetra_ab %>% 
  mutate(PA = ifelse(vivos_n == 0, 0, 1))

tetra_p <- tetra_ab %>% 
  filter(vivos_n > 0)

rm(ordem)

hist(tetra_ab$vivos_n) # Right skewed distribution and zero inflated
```


#### Plotting for exploration
```{r data exploration}

ggplot(tetra_ab, aes(x = site, y = vivos_n, color = subregion)) + 
  geom_boxplot(notch=F) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5))

```

#### VIF and correlations
```{r VIF}

tetra_ab %>% 
  na.omit() %>% 
  dplyr::select(-region,-subregion,-replicate,-site,-easting,-northing, -distance_S, -inclinacao) %>% 
  corvif()

tetra_ab %>% 
  dplyr::select(-region,-subregion,-site, -replicate, -easting,-northing, -distance_S, -inclinacao) %>% 
  PerformanceAnalytics::chart.Correlation(histogram=TRUE, pch=19)

# visual inspection
tetra_ab %>% 
  dplyr::select(-region,-subregion,-replicate,-easting,-northing, -distance_S, -inclinacao, -PA) %>% 
  group_by(site) %>% 
  summarise_if(is.numeric, mean, na.rm = TRUE) %>% 
  gather(key='fatores', value='valores', -site, -vivos_n) %>% 
  ggplot() + 
  geom_point(aes(x= valores, y=vivos_n)) + 
  geom_smooth(aes(x= valores, y=vivos_n)) + 
  facet_wrap(~ fatores, ncol=3, scales = 'free') +
  theme_classic()

```

#### Selecting random structure of presence predictive model
```{r distribution PA}

pa.g.full = glmmTMB(PA ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab + (1|region/subregion/site), data=tetra_ab, REML=T, family = binomial, ziformula=~0) ## random = region, subregion and site ##

pa.g.no_loc = glmmTMB(PA ~  sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab + (1|region/subregion), data=tetra_ab, REML=T, family = binomial, ziformula=~0) ## random = region and subregion ##

pa.g.no_sub = glmmTMB(PA ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab + (1|region/site), data=tetra_ab, REML=T, family = binomial, ziformula=~0) ## random = region and site ##

pa.g.no_reg = glmmTMB(PA ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab + (1|subregion/site), data=tetra_ab, REML=T, family = binomial, ziformula=~0) ## random = subregion and site ##

pa.g.reg = glmmTMB(PA ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab + (1|region), data=tetra_ab, REML=T, family = binomial, ziformula=~0) ## random = only region ##

pa.g.sub = glmmTMB(PA ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab + (1|subregion), data=tetra_ab, REML=T, family = binomial, ziformula=~0) ## random = only subregion ##

pa.g.loc = glmmTMB(PA ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab + (1|site), data=tetra_ab, REML=T, family = binomial, ziformula=~0) ## random = only site ##

pa.none = glmmTMB(PA ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab, data=tetra_ab, REML=T, family = binomial, ziformula=~0) ## random = none ##

cand.set.pa3 <- list(pa.g.full, pa.g.no_reg, pa.g.no_sub, pa.g.no_loc, pa.g.reg, pa.g.sub, pa.g.loc, pa.none)
modnames.pa3 <- c("pa.g.full","pa.g.no_reg","pa.g.no_sub", "pa.g.no_loc","pa.g.reg", "pa.g.sub", "pa.g.loc", "pa.none")
AICctab(cand.set.pa3, mnames= modnames.pa3, base=TRUE, weights=TRUE, logLik=TRUE)

# site level is the best

```

#### Selecting fixed structure of PA model (method = ML)  
```{r fixed PA}

n0.ML = glmmTMB(PA ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + 
                  str_ab + (1|site), data=tetra_ab, 
                REML= FALSE, family=binomial)
summary(n0.ML)

n0.ML2 <- update(n0.ML, .~.-fwd_mean)
summary(n0.ML2)

n0.ML3 <- update(n0.ML2, .~.-str_ab)
summary(n0.ML3)

n0.ML4 <- update(n0.ML3, .~.-rugosity)
summary(n0.ML4)

n0.ML5 <- update(n0.ML4, .~.-incl_graus_tetra)
summary(n0.ML5)

n0.ML6 <- update(n0.ML5, .~.-wf_log)
summary(n0.ML6)

n0.ML7 <- update(n0.ML6, .~.-chl_mean)
summary(n0.ML7)

```


#### Refit with REML and summary 
```{r refit PA}

rm(list=setdiff(ls(), c("tetra_ab", "tetra_p")))

pa.b.REML = glmmTMB(PA ~ sst_mean + (1|site), data=tetra_ab, REML=T, family = binomial)

summary(pa.b.REML)

```

#### Checking model assumptions 
```{r PA REML}

simulateResiduals(pa.b.REML) %>% plot() 
simulateResiduals(pa.b.REML) %>% testDispersion()
simulateResiduals(pa.b.REML) %>% testUniformity()
simulateResiduals(pa.b.REML) %>% plotResiduals(tetra_ab$sst_mean) # some deviation in lower quantile

## check spatial autocorrelation
tetra_ab$easting2  = jitter(tetra_ab$easting)
tetra_ab$northing2 = jitter(tetra_ab$northing)

simulateResiduals(pa.b.REML) %>% 
  testSpatialAutocorrelation(x = tetra_ab$easting2, y = tetra_ab$northing2, plot = T) 
# spatial autocorrelation detected

# test again grouping residuals per site (https://www.rdocumentation.org/packages/DHARMa/versions/0.4.1/topics/testSpatialAutocorrelation)

# calculate simulated residuals per site 
coords <- tetra_ab %>% 
  select(site, easting, northing) %>%
  distinct()

simulateResiduals(pa.b.REML) %>% 
  recalculateResiduals(group = tetra_ab$site) %>% 
  testSpatialAutocorrelation(x = coords$easting, y = coords$northing, plot = T)

# no spatial autocorrelation at site level

```



#### Checking presence data distribution
```{r distribution presence}

dens.nb1 = glmmTMB(vivos_n ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + 
                     str_ab + (1|region/subregion/site), data=tetra_p, REML=T, family = nbinom1)

dens.nb2 = glmmTMB(vivos_n ~  sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + 
                     str_ab + (1|region/subregion/site), data=tetra_p, REML=T, family = nbinom2)

dens.p =   glmmTMB(vivos_n ~  sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + 
                     str_ab + (1|region/subregion/site), data=tetra_p, REML=T, family = poisson)

dens.g =   glmmTMB(vivos_n ~  sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + 
                     str_ab + (1|region/subregion/site), data=tetra_p, REML=T, family = gaussian)

dens.gm =   glmmTMB(vivos_n ~  sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + 
                     str_ab + (1|region/subregion/site), data=tetra_p, REML=T, family = Gamma)

cand.set.dist <- list(dens.nb1, dens.nb2, dens.p, dens.g, dens.gm) ## list of models unedr comparison ##
modnames.dist <- c("dens.nb1","dens.nb2","dens.p", "dens.g", "dens.gm") ## naming different moels ##
AICctab(cand.set.dist, mnames= modnames.dist, base=TRUE, weights=TRUE, logLik=TRUE) ## AICc tab from package bbmle ##

## CONCLUSION: nbinom1 = best distribution ##
```


#### Selecting random structure of presence predictive model
```{r random}

dens.g.full = glmmTMB(vivos_n ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab + (1|region/subregion/site), data=tetra_p, REML=T, family = nbinom1, ziformula=~0) ## random = region, subregion and site ##

dens.g.no_loc = glmmTMB(vivos_n ~  sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab + (1|region/subregion), data=tetra_p, REML=T, family = nbinom1, ziformula=~0) ## random = region and subregion ##

dens.g.no_sub = glmmTMB(vivos_n ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab + (1|region/site), data=tetra_p, REML=T, family = nbinom1, ziformula=~0) ## random = region and site ##

dens.g.no_reg = glmmTMB(vivos_n ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab + (1|subregion/site), data=tetra_p, REML=T, family = nbinom1, ziformula=~0) ## random = subregion and site ##

dens.g.reg = glmmTMB(vivos_n ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab + (1|region), data=tetra_p, REML=T, family = nbinom1, ziformula=~0) ## random = only region ##

dens.g.sub = glmmTMB(vivos_n ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab + (1|subregion), data=tetra_p, REML=T, family = nbinom1, ziformula=~0) ## random = only subregion ##

dens.g.loc = glmmTMB(vivos_n ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab + (1|site), data=tetra_p, REML=T, family = nbinom1, ziformula=~0) ## random = only site ##

dens.none = glmmTMB(vivos_n ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + str_ab, data=tetra_p, REML=T, family = nbinom1, ziformula=~0) ## random = none ##

cand.set.dens3 <- list(dens.g.full, dens.g.no_reg, dens.g.no_sub, dens.g.no_loc, dens.g.reg, dens.g.sub, dens.g.loc, dens.none)
modnames.dens3 <- c("dens.g.full","dens.g.no_reg","dens.g.no_sub", "dens.g.no_loc","dens.g.reg", "dens.g.sub", "dens.g.loc", "dens.none")
AICctab(cand.set.dens3, mnames= modnames.dens3, base=TRUE, weights=TRUE, logLik=TRUE)

## CONCLUSION: Best random structure dens.g.loc (only site) ##
```

#### Selecting fixed structure of predictive model (method = ML)  
```{r fixed presence}

dens.g.ML = glmmTMB(vivos_n ~ sst_mean + chl_mean + fwd_mean + wf_log + incl_graus_tetra + rugosity + 
                      str_ab + (1|site), data=tetra_p, REML=FALSE, family = nbinom1, ziformula=~0)
summary(dens.g.ML)

dens.g.ML1 <- update(dens.g.ML, .~.-str_ab)
summary(dens.g.ML1)

dens.g.ML2 <- update(dens.g.ML1, .~.-chl_mean)
summary(dens.g.ML2)

dens.g.ML3 <- update(dens.g.ML2, .~.-fwd_mean)
summary(dens.g.ML3)

dens.g.ML4 <- update(dens.g.ML3, .~.-rugosity)
summary(dens.g.ML4)

dens.g.ML5 <- update(dens.g.ML4, .~.-incl_graus_tetra)
summary(dens.g.ML5)

dens.g.ML6 <- update(dens.g.ML5, .~.-wf_log)
summary(dens.g.ML6)

## CONCLUSION: selected model includes: only SST ##
```

#### Refit with REML and summary 
```{r REML}

rm(list=setdiff(ls(), c("tetra_ab", "tetra_p", "coords")))

dens.nb1.REML = glmmTMB(vivos_n ~ sst_mean + (1|site), data=tetra_p, REML=T, family = nbinom1, ziformula=~0)
summary(dens.nb1.REML)

## dispersion parameter
# https://github.com/glmmTMB/glmmTMB/issues/224
dispfun <- function(m) {
  r <- residuals(m, type="pearson")
  n <- df.residual(m)
  dsq <- sum(r^2)
  c(dsq = dsq, n = n, disp = dsq/n)
}

dispfun(dens.nb1.REML) #OK

```

#### Checking model assumptions 
```{r residuals}

simulateResiduals(dens.nb1.REML) %>% plot() ## quantiles show deviation 
simulateResiduals(dens.nb1.REML) %>% testOutliers()   ## OK 
simulateResiduals(dens.nb1.REML) %>% testDispersion() ## OK
simulateResiduals(dens.nb1.REML) %>% testUniformity() ## OK
simulateResiduals(dens.nb1.REML) %>% plotResiduals(tetra_p$sst_mean, xlab = "SST mean", ylab = "scaled residuals")
simulateResiduals(dens.nb1.REML) %>% hist()

## CONCLUSION: simulated residuals look fine ##

### checking residuals against all variables
tetra_p %>% 
  filter(!is.na(vivos_n)) %>% 
  dplyr::select(site, chl_mean, sst_mean, fwd_mean, wf_log, rugosity, inclinacao, str_ab, str_size, vivos_n) %>% 
  pivot_longer(names_to = 'fatores', values_to ='valores', cols=chl_mean:str_size) %>% 
  aggregate(.~site+fatores, FUN=mean, data=.) %>% 
  ggplot() + 
  geom_point(aes(x= valores, y=vivos_n)) + 
  geom_smooth(aes(x= valores, y=vivos_n)) +
  facet_wrap(~ fatores, ncol=3, scales = 'free') +
  theme_classic()

```

#### Plotting to check spatial autocorrelation
```{r check spatial}

## Creating jittered values for coordinates since testing spatial autocorrelation in DHARMa only works for unique values ##
tetra_p$easting2  = jitter(tetra_p$easting)
tetra_p$northing2 = jitter(tetra_p$northing)

simulateResiduals(dens.nb1.REML) %>% testSpatialAutocorrelation(x = tetra_p$easting2, y = tetra_p$northing2, plot = T) 
# evidence of spatial autocorrelation

# calculate simulated residuals per site 
simulateResiduals(dens.nb1.REML) %>% 
  recalculateResiduals(group = tetra_p$site) %>% 
  testSpatialAutocorrelation(x = coords$easting, y = coords$northing, plot = T)

## no spatial correlation at the site level ##

```


#### Fitting spatial models
```{r spatial model}

# INLA models were ambiguous

```


#### Fitting GAM models

```{r gam}

library(mgcv)
library(scam)

## reducing number of samples to site level to decrease dispersion
tetra1 <- left_join(tetra_ab %>% select(-str_ab, -replicate, -subregion) %>% aggregate(vivos_n~site+region, ., mean),
                    tetra_ab %>% select(-str_ab, -replicate, -subregion) %>% aggregate(cbind(sst_mean, easting, northing, fwd_mean)~site+region, ., mean))

## check relationship shape
xyplot(tetra1$vivos_n ~ tetra1$sst_mean,type=c("p", "smooth"), pch=21, cex=1.2, col="black", fill=c("light blue")) # sigmoid-like shape

# comparing GAM with bs spline and SCAM with constrained shape (sigmoid)
gam_lp = gam(vivos_n ~ s(sst_mean), data=tetra1)
scam_lp = scam::scam(vivos_n ~ s(sst_mean, bs="mpd"), data=tetra1)

summary(gam_lp)
summary(scam_lp)

gam.check(gam_lp)
gam.check(scam_lp)

plot(gam_lp, las=1)
plot(scam_lp, las=1)

AIC(scam_lp, gam_lp) # SCAM had lower AIC


simulateResiduals(scam_lp) %>% plot() ## OK
simulateResiduals(scam_lp) %>% testDispersion() ## OK
simulateResiduals(scam_lp) %>% testUniformity() ## OK
simulateResiduals(scam_lp) %>% plotResiduals(tetra1$sst_mean, xlab = "SST mean", ylab = "scaled residuals") ## some deviance in lower quantile

## check spatial autocorrelation
simulateResiduals(scam_lp) %>% 
  testSpatialAutocorrelation(x = tetra1$easting, y = tetra1$northing, plot = T) 
# no autocorrelation

```


#### Predictions of best model
```{r predictions}

## sst_mean

sst_pred <- predict(scam_lp, se=T) %>% # check if this is the right way to predict values
  data.frame() %>% 
  mutate(sst_mean = tetra1$sst_mean) %>% 
  ggplot(aes(x = sst_mean, y = fit)) +
    geom_point(size= 2.2, alpha=0.5) + 
    geom_ribbon(aes(y = NULL, ymin = fit-se.fit, ymax = fit+se.fit), fill = "grey50", alpha=0.5)+
    geom_line(mapping = aes(x = sst_mean, y = fit), size = 1.3, color="black") +
    theme_classic() + theme(axis.text.x = element_text(size=12, color="black"), 
                            axis.text.y = element_text(size=12, color="black"),
                            axis.title.x = element_text(color="black", size=14),
                            axis.title.y = element_text(color="black", size=16)) +
    labs(x = expression(paste("SST (",~degree~C,")")), 
         y = expression(paste("Density ", "(", ind.100cm^-2, ")"))) +
    theme(strip.text.x = element_text(size = 10), strip.text.y = element_text(size = 10))+
    theme(axis.ticks = element_line(colour = "black", size = 0.8)) +
    theme(axis.line = element_line(colour = 'black', size = 0.8)) +
    theme(axis.ticks.length = unit(4, "pt"), legend.position = "")

ggsave("sst_pred.tiff", dpi = 600, compression = 'lzw')

```
